---
title: "Formatting_MG"
author: "Megan Glass"
date: "2023-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggfootball)
library(forecast)
library(caret)
```

```{r}
tackles <- data.table::fread("tackles.csv")

plays <- data.table::fread("plays.csv")
```

## Preprocess Data

```{r}
heights_to_inches <- function(height_strings) {
  # Initialize an empty vector to store the results
  result <- numeric(length(height_strings))
  
  # Loop through each height string in the vector
  for (i in seq_along(height_strings)) {
    # Split the height string into feet and inches
    height_components <- strsplit(height_strings[i], "-")[[1]]
    
    # Convert feet and inches to inches
    feet <- as.numeric(height_components[1])
    inches <- as.numeric(height_components[2])
    
    # Calculate total inches and store the result
    result[i] <- feet * 12 + inches
  }
  
  return(result)
}
```




# Derived Variables 

want player with the ball and player who made the tackle (done)
use missed tackle (done)

## Merge with Plays

Get ballCarrierID from plays

```{r}
week6 <- data.table::fread("tackle_events/tackle_events_week6.csv")
week6 <- week6[, -24]
colnames(week6)[5] <- "displayName"
week6$height_in <- heights_to_inches(week6$height)
play373 <- week6 %>%
  filter(gameId == 2022101605, playId == 373)

football_loc <- play373 %>%
  filter(displayName == "football")
    ## line of scrimmage
    line_of_scrimmage <- football_loc$x[1]
    ## ball_start
    ball_start <- football_loc[1,]

play373$ballCarrierIndicator <- 0

ballCarrierId <- plays %>%
  filter(gameId == 2022101605, playId == 373) %>%
  select(ballCarrierId)

play373$ballCarrierIndicator[which(play373$nflId == ballCarrierId$ballCarrierId)] <- 1

table(play373$ballCarrierIndicator)
```

## Merge with tackle (attempt)
get who made the tackle from plays
```{r}
play373$tackleIndicator <- 0

tackleId <- tackles %>%
  filter(gameId == 2022101605, playId == 373) %>%
  select(nflId)

play373$tackleIndicator[which(play373$nflId == tackleId$nflId)] <- 1

table(play373$tackleIndicator)
```

points of time:
snap 
player crosses line of scrimmage
tackle

variables to include for each time:
orientation
time from snap to tackle
speeds directions

## Distance from the football (at the time of the snap)
```{r}
players <- unique(play373$displayName)
frames <- unique(play373$frameId)

play373$football_dist_snap <- NA
play373$orientation_snap <- NA
play373$snap_tackle_time <- NA
play373$speed_diff_bc_vs_tackle_snap <- NA

snap_loc_x <- ball_start$x
snap_loc_y <- ball_start$y
snap_time <-  ball_start$time
snap_frame <- ball_start$frameId

tackle_time <- play373 %>%
  filter(tackleIndicator == 1, event == "tackle") %>%
  select(time)

time_diff <- tackle_time$time - snap_time

player_x <- play373 %>% 
  filter(tackleIndicator ==1, frameId == snap_frame) %>%
  select(x)

player_y <- play373 %>% 
  filter(tackleIndicator ==1, frameId == snap_frame) %>%
  select(y)

bc_speed <- play373 %>%
  filter(ballCarrierIndicator == 1, frameId ==snap_frame) %>%
  select(s)

tackle_speed <- play373 %>%
  filter(tackleIndicator == 1, frameId==snap_frame) %>%
  select(s)

speed_diff<- bc_speed$s - tackle_speed$s

orientation <- play373 %>%
  filter(tackleIndicator == 1, frameId == snap_frame) %>%
  select(o)

play373$football_dist_snap[which(play373$tackleIndicator==1)] <- sqrt((snap_loc_x - player_x)^2+(snap_loc_y - player_y)^2)$x
    
play373$orientation_snap[which(play373$tackleIndicator == 1)] <- orientation$o

play373$snap_tackle_time[which(play373$tackleIndicator ==1)] <- time_diff
  
play373$speed_diff_bc_vs_tackle_snap[which(play373$tackleIndicator == 1)] <- speed_diff

play373 %>%
  filter(tackleIndicator == 1)
```

## Distance from the football (at the tackle)
```{r}
play373$football_dist_t <- NA
play373$orientation_t <- NA
play373$speed_diff_bc_vs_tackle_t <- NA


tackle_x <- play373 %>%
  filter(event == "tackle", displayName == "football") %>% 
  select(x)
tackle_y <- play373 %>%
  filter(event == "tackle", displayName == "football") %>% 
  select(y)

tackle_frame <- play373 %>%
  filter(event == "tackle", displayName == "football") %>%
  select(frameId)

player_x <- play373 %>% 
  filter(tackleIndicator ==1, frameId == tackle_frame$frameId) %>%
  select(x)

player_y <- play373 %>% 
  filter(tackleIndicator ==1, frameId == tackle_frame$frameId) %>%
  select(y)

bc_speed <- play373 %>%
  filter(ballCarrierIndicator == 1, frameId ==tackle_frame$frameId) %>%
  select(s)

tackle_speed <- play373 %>%
  filter(tackleIndicator == 1, frameId ==tackle_frame$frameId) %>%
  select(s)

speed_diff<- bc_speed$s[1] - tackle_speed$s[1]

orientation <- play373 %>%
  filter(tackleIndicator == 1, frameId == tackle_frame$frameId) %>%
  select(o) %>%
  .[1]

play373$football_dist_t[which(play373$tackleIndicator == 1)] <- sqrt((tackle_x$x - player_x$x)^2+(tackle_y$y - player_y$y)^2)
play373$orientation_t[which(play373$tackleIndicator == 1)] <- orientation$o
play373$speed_diff_bc_vs_tackle_t[which(play373$tackleIndicator == 1)] <- speed_diff
play373 %>%
  filter(tackleIndicator == 1)
```

## Variables at the time of the snap
```{r}
play373$football_dist <- NA

for (i in 1:length(players)){
  for (j in 1:length(frames)){
    football_x <- play373$x[which(play373$displayName == "football" &
                                    play373$frameId == frames[j])]
    football_y <- play373$y[which(play373$displayName == "football" &
                                    play373$frameId == frames[j])]
    player_x <- play373$x[which(play373$displayName == players[i] &
                                    play373$frameId == frames[j])]
    player_y <- play373$y[which(play373$displayName == players[i] &
                                    play373$frameId == frames[j])]
    
    dist <- sqrt((football_x - player_x)^2+(football_y - player_y)^2)
    
    play373$football_dist[which(play373$displayName == players[i] &
                                    play373$frameId == frames[j])] <- dist
    
  }
}
```





## Get the one row of tackle

```{r}
the_tackle <- tackles %>%
  filter(gameId == 2022101605, playId == 373)


tackle_player <- play373 %>%
  filter(tackleIndicator == 1, event == "tackle") 

merge(the_tackle, tackle_player)
```



### All weeks data
```{r}
file_names <- c("tackle_events/tackle_events_week1.csv", 
                "tackle_events/tackle_events_week2.csv",
                "tackle_events/tackle_events_week3.csv",
                "tackle_events/tackle_events_week4.csv",
                "tackle_events/tackle_events_week5.csv",
                "tackle_events/tackle_events_week6.csv",
                "tackle_events/tackle_events_week7.csv",
                "tackle_events/tackle_events_week8.csv",
                "tackle_events/tackle_events_week9.csv")

# Initialize an empty data frame
all_week_data <- data.frame()

# Loop through the files and read data into the combined_data data frame
for (file in file_names) {
  current_data <- data.table::fread(file, header = TRUE)  # Adjust parameters based on your CSV files
  
  current_data <- current_data[, -24]
  colnames(current_data)[5] <- "displayName"
  current_data$height_in <- heights_to_inches(current_data$height)
  current_data <- current_data[,-c(8,21,22,19)]

  all_week_data <- bind_rows(all_week_data, current_data)
}

```

```{r}
# not this
all_week_data <- data.table::fread("tackle_events/tackle_events_week1.csv")
all_week_data <- all_week_data[, -24]
  colnames(all_week_data)[5] <- "displayName"
  all_week_data$height_in <- heights_to_inches(all_week_data$height)
```



### All weeks
```{r}
test.games <- unique(all_week_data$gameId)
selected.game <- test.games[1]

all_week_data %>%
  filter(gameId == selected.game) %>%
  group_by(playId) %>%
  summarise(count = n()) 


tackles %>%
  filter(gameId == selected.game) %>%
  group_by(playId) %>%
  summarise(count = n())
```


```{r}
week.games <- unique(all_week_data$gameId)

week.tackles <- data.frame()

for(g in 1:length(week.games)) {
  selected.game <- week.games[g]

  game <- all_week_data %>%
    filter(gameId == selected.game) 
  
  game.plays <- unique(game$playId)
  
  game$ballCarrierIndicator <- 0
  game$tackleIndicator <- 0
  game$football_dist_snap <- NA
  game$orientation_snap <- NA
  game$snap_tackle_time <- NA
  game$speed_diff_bc_vs_tackle_snap <- NA
  game$football_dist_t <- NA
  game$orientation_t <- NA
  game$speed_diff_bc_vs_tackle_t <- NA
  game$player_start_x <- NA
  game$player_start_y <- NA
  game$x_dist_football <- NA
  game$y_dist_football <- NA
  game$ball_start_x <- NA
  game$ball_start_y <- NA
  
  for (i in 1:length(game.plays)) {
    ## football locations
    football_loc <- game %>%
      filter(playId == game.plays[i], displayName == "football")
    ## line of scrimmage
    line_of_scrimmage <- football_loc$x[1]
    ## ball_start
    ball_start <- football_loc[1,]
    
    snap_loc_x <- ball_start$x
    snap_loc_y <- ball_start$y
    snap_time <-  ball_start$time
    snap_frame <- ball_start$frameId
  
    ballCarrierId <- plays %>%
      filter(gameId == selected.game, playId == game.plays[i])%>%
      select(ballCarrierId)
  
    game$ballCarrierIndicator[which(game$nflId == ballCarrierId$ballCarrierId &
                                      game$playId == game.plays[i])] <- 1
    bc_speed_snap <- game %>%
      filter(playId == game.plays[i],ballCarrierIndicator == 1, frameId ==snap_frame) %>%
      select(s)
    
    tackleId <- tackles %>%
      filter(gameId == selected.game, playId == game.plays[i]) %>%
      select(nflId)
  
    
    for(j in 1:length(tackleId)){
      game$tackleIndicator[which(game$nflId == tackleId$nflId[j] &
                                      game$playId == game.plays[i])] <- 1
      
      tackle_time <- game %>%
        filter(playId == game.plays[i], tackleIndicator == 1, event == "tackle", nflId == tackleId$nflId[j]) %>%
        select(time)
      
      time_diff_snap <- difftime(tackle_time$time, snap_time)
      
      game$ball_start_x[which(game$tackleIndicator==1 &
                                        game$playId == game.plays[i] &
                                        game$nflId == tackleId$nflId[j])] <- snap_loc_x
       game$ball_start_y[which(game$tackleIndicator==1 &
                                        game$playId == game.plays[i] &
                                        game$nflId == tackleId$nflId[j])] <- snap_loc_y
      
      player_x <- game %>%
      filter(playId == game.plays[i], tackleIndicator ==1, frameId == snap_frame, nflId == tackleId$nflId[j]) %>%
        select(x)
      
      game$player_start_x[which(game$tackleIndicator==1 &
                                        game$playId == game.plays[i] &
                                        game$nflId == tackleId$nflId[j])] <- player_x
      game$x_dist_football[which(game$tackleIndicator==1 &
                                        game$playId == game.plays[i] &
                                        game$nflId == tackleId$nflId[j])] <- snap_loc_x - player_x
      
      player_y <- game %>%
        filter(playId == game.plays[i],tackleIndicator ==1, frameId == snap_frame, nflId == tackleId$nflId[j]) %>%
        select(y)
      game$player_start_y[which(game$tackleIndicator==1 &
                                        game$playId == game.plays[i] &
                                        game$nflId == tackleId$nflId[j])] <- player_y
      game$y_dist_football[which(game$tackleIndicator==1 &
                                        game$playId == game.plays[i] &
                                        game$nflId == tackleId$nflId[j])] <- snap_loc_y - player_y
      
      tackle_speed_snap <- game %>%
        filter(playId == game.plays[i],tackleIndicator == 1, frameId==snap_frame, nflId == tackleId$nflId[j]) %>%
        select(s)
      
      speed_diff_snap<- bc_speed_snap$s - tackle_speed_snap$s
      
      orientation_snap <- game %>%
        filter(playId == game.plays[i],tackleIndicator == 1, frameId == snap_frame, nflId == tackleId$nflId[j]) %>%
        select(o)
      
      game$football_dist_snap[which(game$tackleIndicator==1 &
                                        game$playId == game.plays[i] &
                                        game$nflId == tackleId$nflId[j])] <- sqrt((snap_loc_x - player_x)^2+(snap_loc_y - player_y)^2)$x
        
    game$orientation_snap[which(game$tackleIndicator==1 &
                                        game$playId == game.plays[i] &
                                        game$nflId == tackleId$nflId[j])] <- orientation_snap$o
    
    game$snap_tackle_time[which(game$tackleIndicator==1 &
                                        game$playId == game.plays[i] &
                                        game$nflId == tackleId$nflId[j])]<- time_diff_snap
      
    game$speed_diff_bc_vs_tackle_snap[which(game$tackleIndicator==1 &
                                        game$playId == game.plays[i] &
                                        game$nflId == tackleId$nflId[j])] <- speed_diff_snap
      tackle_x <- game %>%
        filter(playId == game.plays[i], event == "tackle", displayName == "football") %>% 
        select(x)
      tackle_y <- game %>%
        filter(playId == game.plays[i], event == "tackle", displayName == "football") %>% 
        select(y)
      
      tackle_frame <- game %>%
        filter(playId == game.plays[i], event == "tackle", displayName == "football") %>%
        select(frameId)
      bc_speed_t <- game %>%
        filter(playId == game.plays[i], ballCarrierIndicator == 1, frameId ==tackle_frame$frameId) %>%
        select(s)
    
      player_x_t <- game %>%
        filter(playId == game.plays[i], tackleIndicator ==1, frameId == tackle_frame$frameId, nflId == tackleId$nflId[j]) %>%
        select(x)
      
      player_y_t <- game %>%
        filter(playId == game.plays[i], tackleIndicator ==1, frameId == tackle_frame$frameId, nflId == tackleId$nflId[j]) %>%
        select(y)
      
      tackle_speed_t <- game %>%
        filter(playId == game.plays[i], tackleIndicator == 1, frameId ==tackle_frame$frameId, nflId == tackleId$nflId[j]) %>%
        select(s)
      
      speed_diff_t<- bc_speed_t$s[1] - tackle_speed_t$s[1]
      
      orientation_t <- game %>%
        filter(playId == game.plays[i], tackleIndicator == 1, frameId == tackle_frame$frameId, nflId == tackleId$nflId[j]) %>%
        select(o) %>%
        .[1]
      
      game$football_dist_t[which(game$tackleIndicator == 1&
                                        game$playId == game.plays[i] &
                                        game$nflId == tackleId$nflId[j])] <- sqrt((tackle_x$x - player_x_t$x)^2+(tackle_y$y - player_y_t$y)^2)
      game$orientation_t[which(game$tackleIndicator == 1&
                                          game$playId == game.plays[i] &
                                          game$nflId == tackleId$nflId[j])] <- orientation_t$o
      game$speed_diff_bc_vs_tackle_t[which(game$tackleIndicator == 1 &
                                          game$playId == game.plays[i] &
                                          game$nflId == tackleId$nflId[j])] <- speed_diff_t
    }
    
  }

  the_tackle <- tackles %>%
    filter(gameId == selected.game, playId %in% game.plays)
  
  
  tackle_player <- game %>%
    filter(tackleIndicator == 1, event == "tackle") 
  
  week.tackles <- rbind(week.tackles, merge(the_tackle, tackle_player, by = c("gameId", "playId", "nflId")))

}
```

```{r}
game %>%
  filter(playId == game.plays[i], event == "tackle") %>%
        select(frameId)
```



```{r}
data.table::fwrite(week.tackles, file = "tackle_events/all_week_tackle_data.csv")
```

